---
title: "Multivariate Data Analysis - Presentation"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install Packages
list.of.packages <- c("data.table"
                      ,"car"
                      ,"ggcorrplot"
                      ,"ggthemes"
                      ,"gridExtra"
                      ,"ggplot2"
                      )
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, dependencies = T)

# Load Packages
library(data.table)
library(ggplot2)
library(gridExtra)
library(ggthemes)
library(ggcorrplot)
library(car)
```



## Step 1: Download the dataset from Kaggle

To simplify the process, I am using my github account to host the dataset but the original dataset is available here:
[Kaggle: House Sales in King County](https://www.kaggle.com/harlfoxem/housesalesprediction)
```{r}
wd <- getwd()
filep <- "https://raw.githubusercontent.com/anuj-kapil/mda/master/Data/kc_house_data.csv"
filename <- "kc_house_data.csv"
download.file(url=filep, destfile=filename)
```

## Step 2: Read the downloaded file to memory

We are using the data.table package from R to read the file in to memory

```{r}
kc_houses<-fread('kc_house_data.csv')
```

## Step 3: Exploratory Analysis

Let's have a look at the number of features and observations in the dataset
```{r}
# Rows and columns
dim(kc_houses)
```

Quick look at the data types and ranges of data. Below summary shows that there are no null or missing values
```{r}
# No missing values
summary(kc_houses)
```
The dates are loaded as string of characters. Let's convert them to look at the range of values. The summary now shows that the dates are in the range of May 2014 - May 2015
```{r}
# Convert Dates
kc_houses$date<-as.IDate(kc_houses$date, format = "%m/%d/%Y")
summary(kc_houses)
```